---
title: "Data_Processing"
author: "Team"
date: "April 18, 2016"
output: html_document
---
#EVENTMSGTYPE
1 - Make 2 - Miss 3 - Free Throw 4 - Rebound 5 - out of bounds / Turnover / Steal 6 - Personal Foul 7 - Violation 8 - Substitution 9 - Timeout 10 - Jumpball 12 - Start Q1? 13 - Start Q2?

#EVENTMSGACTIONTYPE
1 - Jumpshot 2 - Lost ball Turnover 3 - ? 4 - Traveling Turnover / Off Foul 5 - Layup 7 - Dunk 10 - Free throw 1-1 11 - Free throw 1-2 12 - Free throw 2-2 40 - out of bounds 41 - Block/Steal 42 - Driving Layup 50 - Running Dunk 52 - Alley Oop Dunk 55 - Hook Shot 57 - Driving Hook Shot 58 - Turnaround hook shot 66 - Jump Bank Shot 71 - Finger Roll Layup 72 - Putback Layup 108 - Cutting Dunk Shot



```{r setup, include=FALSE}
setwd("/Users/bobminnich/Documents/Columbia/Courses/Applied_Data_Science/FinalProject/finalproject-p5-team3")
library(dplyr)
library(ggplot2)
load("teamdata.Rdata")
teamdata$EVENTMSGTYPE = as.numeric(teamdata$EVENTMSGTYPE)
teamdata$GAME_ID = as.factor(as.character(teamdata$GAME_ID))
teamdata$EVENTMSGTYPE = as.numeric(teamdata$EVENTMSGTYPE)
teamdata$HOMEDESCRIPTION = tolower(as.character(teamdata$HOMEDESCRIPTION))
teamdata$VISITORDESCRIPTION = tolower(as.character(teamdata$VISITORDESCRIPTION))
teamdata$PERIOD = as.numeric(teamdata$PERIOD)
teamdata$EVENTNUM = (as.numeric(teamdata$EVENTNUM))
teamdata$WCTIMESTRING = as.character(teamdata$WCTIMESTRING)
teamdata$PCTIMESTRING = as.character(teamdata$PCTIMESTRING)

teamdata = teamdata[!duplicated(teamdata),] 
# 
# df_fouls = filter(teamdata, EVENTMSGTYPE=="6" ) %>% group_by(GAME_ID) %>% summarise(count = n())
# #0021400020 had the most fouls Thunder vs Clippers
# which.max(df_fouls$count)
# ggplot(df_fouls, aes(x=count)) +
#     geom_histogram(binwidth=1, alpha=.5, position="identity")
# 



teamdata$HFoulName = "NULL"
teamdata$HMake = "NULL"
teamdata$HFreeName = "NULL"
teamdata$HTake = 0

teamdata$VFoulName = "NULL"
teamdata$VMake = "NULL"
teamdata$VFreeName = "NULL"
teamdata$VTake = 0

teamdata$min <- lapply(strsplit(teamdata$PCTIMESTRING, ":"), "[", 1)
teamdata$sec <- lapply(strsplit(teamdata$PCTIMESTRING, ":"), "[", 2)
teamdata$min4 <- floor(as.numeric(teamdata$min)/4)
teamdata$min2 <- floor(as.numeric(teamdata$min)/2)
teamdata$Time <- (as.numeric(teamdata$PERIOD) - 1)*12*60 +abs((as.numeric(teamdata$min)-12)*60 - (0-as.numeric(teamdata$sec)))

for(i in 1:nrow(teamdata)){
  if(grepl("foul", teamdata$HOMEDESCRIPTION[i])){
    if(grepl("take", teamdata$HOMEDESCRIPTION[i])){
      teamdata$HFoulName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[1]
      teamdata$HTake[i] =1
      teamdata$HMake[i] = "NULL"}
    else{
      teamdata$HFoulName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[1]
      teamdata$HMake[i] = "NULL"
      }
  }
    else if(grepl("miss", teamdata$HOMEDESCRIPTION[i])){
    teamdata$HFreeName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[2]
    teamdata$HMake[i] = 0
}
  else if(grepl("free", teamdata$HOMEDESCRIPTION[i])){
    teamdata$HFreeName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[1]
    teamdata$HMake[i] = 1
  }else{
    teamdata$HFreeName[i] = "NULL"
    teamdata$HFoulName[i] = "NULL"
    teamdata$HMake[i] = "NULL"
  }
  
  if(grepl("foul", teamdata$VISITORDESCRIPTION[i])){
    if(grepl("take", teamdata$VISITORDESCRIPTION[i])){
      teamdata$VFoulName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[1]
      teamdata$VTake[i] =1
      teamdata$VMake[i] = "NULL"}
    else{
      teamdata$VFoulName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[1]
      teamdata$VMake[i] = "NULL"
      }
  }
    else if(grepl("miss", teamdata$VISITORDESCRIPTION[i])){
    teamdata$VFreeName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[2]
    teamdata$VMake[i] = 0
}
  else if(grepl("free", teamdata$VISITORDESCRIPTION[i])){
    teamdata$VFreeName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[1]
    teamdata$VMake[i] = 1
  }else{
    teamdata$VFreeName[i] = "NULL"
    teamdata$VFoulName[i] = "NULL"
    teamdata$VMake[i] = "NULL"
  }
}


```


#Team Statistics
```{r}
df_per_game = filter(teamdata, EVENTMSGTYPE=="3")%>%group_by(GAME_ID,PERIOD)%>% summarise(count = n())
df_game = filter(teamdata, EVENTMSGTYPE=="3")%>%group_by(GAME_ID)%>% summarise(count = n())

mean(df_game$count)
median(df_game$count)
range(df_game$count)

df_per_game = df_per_game[complete.cases(df_per_game),]
ggplot(df_per_game, aes(x=factor(PERIOD),count)) + geom_boxplot()


ggplot(df_per_game, aes(x = factor(GAME_ID), y = count)) + 
  geom_bar(stat = "identity")+ 
  theme(axis.title.x = element_blank())+ 
  scale_x_discrete(breaks=NULL)
  

```
#Player Statistics
HFreeName

```{r}
df_per_game = filter(teamdata, EVENTMSGTYPE=="3",HFreeName == "jordan" )%>%group_by(GAME_ID,PERIOD)%>% summarise(count = n())
df_game = filter(teamdata, EVENTMSGTYPE=="3",HFreeName == "jordan" )%>%group_by(GAME_ID)%>% summarise(count = n())
df_game = df_game[complete.cases(df_game),]
mean(df_game$count)
median(df_game$count)
range(df_game$count)
quantile(df_game$count)

df_game[df_game$count > 7,]
```


#Player Statistics
Dean
```{r}
df_game = filter(teamdata, EVENTMSGTYPE=="3",HFreeName == "jordan" | VFreeName == "jordan" )%>%group_by(GAME_ID)%>% summarise(count = n())
df_game = df_game[complete.cases(df_game),]
mean(df_game$count)
median(df_game$count)
range(df_game$count)
quantile(df_game$count)

ggplot(df_game, aes(x=count)) +
    geom_histogram(binwidth=.5, colour="red", fill="white") + 
  xlab("Number of Fouls")+
  ylab("Number of Games")

out = df_game[df_game$count>8,]$GAME_ID
```

#Jordan per 4min intervlas
HFreeName

```{r}
df_game = filter(teamdata, EVENTMSGTYPE=="3",HFreeName == "jordan" )%>%group_by(GAME_ID,min4,PERIOD)%>% summarise(count = n())

mean(df_game$count)
median(df_game$count)
range(df_game$count)
quantile(df_game$count)
#0021400802 Spurs
#0021400908 Clippers lost to Blazers
#Games with greater than 75%
hack = unique(df_game[df_game$count>2,]$GAME_ID)

df_game$gre = as.numeric(df_game$count > 2)
df_game$cgre = df_game$count * df_game$gre

out = df_game%>%group_by(GAME_ID)%>% summarise(avg = mean(cgre))
out2 = out[out$avg > 0,]

quantile(out2$avg)
out2$GAME_ID
```

#Filter on game and plot fould per 4 min
```{r}
df_game = filter(teamdata, EVENTMSGTYPE=="3",HFreeName == "jordan" ,GAME_ID == "0021400599")%>%group_by(Time)%>% summarise(count = n())

df_game = filter(teamdata, EVENTMSGTYPE=="3",HFreeName == "jordan" ,GAME_ID == "0021400599")

a = filter(teamdata,GAME_ID == "0021400599")
View(a)
x = 0:(12*60*4)
y = 0:(12*60*4) * 0
y[df_game$Time] = df_game$count
plot(x,y)

p <- ggplot(mry, aes(x=year, y=number, group=rating))
p + geom_line()
```


```{r}
df_game =filter(teamdata,HTake==1)%>%group_by(GAME_ID)%>% summarise(count = n())
quantile(df_game$count)
a = filter(teamdata,HTake==1)%>%group_by(GAME_ID)
ggplot(df_game, aes(x=count)) +
    geom_histogram(binwidth=.5, colour="red", fill="white") + 
  xlab("Number of Take Fouls in one Game")+
  ylab("Number of Games")

quantile(df_game$count)
out2 = df_game[df_game$count >= 5,]$GAME_ID 


```

```{r}
df_game =filter(teamdata, EVENTMSGTYPE=="3",VFreeName == "jordan" ,GAME_ID == "0021400659")%>%group_by(PERIOD,min4)%>% summarise(count = n())

a = filter(teamdata,GAME_ID == "0021400659")

quantile(df_game$count)
a = filter(teamdata,HTake==1)%>%group_by(GAME_ID)
#View(a)
ggplot(df_game, aes(x=count)) +
    geom_histogram(binwidth=.5, colour="black", fill="white")

quantile(df_game$count)
out2 = df_game[df_game$count > 5,]$GAME_ID


```

