---
title: "Data_Processing"
author: "Team"
date: "April 18, 2016"
output: html_document
---
#EVENTMSGTYPE
1 - Make 2 - Miss 3 - Free Throw 4 - Rebound 5 - out of bounds / Turnover / Steal 6 - Personal Foul 7 - Violation 8 - Substitution 9 - Timeout 10 - Jumpball 12 - Start Q1? 13 - Start Q2?

#EVENTMSGACTIONTYPE
1 - Jumpshot 2 - Lost ball Turnover 3 - ? 4 - Traveling Turnover / Off Foul 5 - Layup 7 - Dunk 10 - Free throw 1-1 11 - Free throw 1-2 12 - Free throw 2-2 40 - out of bounds 41 - Block/Steal 42 - Driving Layup 50 - Running Dunk 52 - Alley Oop Dunk 55 - Hook Shot 57 - Driving Hook Shot 58 - Turnaround hook shot 66 - Jump Bank Shot 71 - Finger Roll Layup 72 - Putback Layup 108 - Cutting Dunk Shot



```{r setup, include=FALSE}
setwd("~/Google Drive/2016 Spring/Applied Data Science/Project 5/finalproject-p5-team3Tara")
library(dplyr)
library(ggplot2)
library(lubridate)
load("teamdata.Rdata")
# load("game2011.Rdata")
teamdata$EVENTMSGTYPE = as.numeric(teamdata$EVENTMSGTYPE)
teamdata$GAME_ID = as.factor(as.character(teamdata$GAME_ID))
teamdata$EVENTMSGACTIONTYPE = as.numeric(teamdata$EVENTMSGACTIONTYPE)
teamdata$PERIOD = as.numeric(teamdata$PERIOD)

# Remove duplicates
teamdata <- teamdata[!duplicated(teamdata), ]

# Filter foul counts
df_fouls = filter(teamdata, EVENTMSGTYPE == 6) %>% group_by(GAME_ID) %>% summarise(count = n())
# Game 0021400020 had the most FOULS (Thunder vs Clippers)
# which.max(df_fouls$count)
df_fouls_game = filter(teamdata, GAME_ID=="0021400020" & (EVENTMSGTYPE=="6" | EVENTMSGTYPE=="3"))

# Filter 'Personal Take Foul' counts --> Game 0021400802 has the most PERSONAL TAKE FOULS (Spurs vs Clippers)
df_ptf <- filter(teamdata, EVENTMSGACTIONTYPE == 28) %>% group_by(GAME_ID) %>% summarise(count = n())

```

# SCORE MARGIN AND NUMBER OF FOULS THROUGHOUT GAME

```{r setup, include=FALSE}
# Filter out any game (change ID)
game = filter(teamdata, GAME_ID=="0021400802")

# Data manipulation:
## Fix SCOREMARGIN
game$SCOREMARGIN[1] = '0'
for(i in 1:nrow(game)) {
  if(game$SCOREMARGIN[i] == 'TIE')
    game$SCOREMARGIN[i] <- '0'
  else if(game$SCOREMARGIN[i] == 'NULL')
    game$SCOREMARGIN[i] <- game$SCOREMARGIN[i - 1]
}
game$SCOREMARGIN <- as.numeric(game$SCOREMARGIN)

## Fix SCORE
#game$SCORE[1] = '0 - 2'
#for(i in 1:nrow(game)) {
#  if(game$SCORE[i] == 'NULL')
#    game$SCORE[i] <- game$SCORE[i - 1]
#}

## Add timestamps min:sec left AND sec into game
game$MINSECLEFT <- ms(game$PCTIMESTRING) + minutes((4 - game$PERIOD) * 12)
game$SECINTOGAME <- 2880 - period_to_seconds(game$MINSEC)

## Compute number of fouls in every four minute interval
game$GROUP <- ceiling(game$SECINTOGAME / 240)
# "HOMEDESCRIPTION == NULL" makes sure we are only getting visitor team's fouls. CHANGE AS NECESSARY.
temp <- filter(game, EVENTMSGTYPE == 6 & HOMEDESCRIPTION == 'NULL') %>% group_by(GROUP) %>% summarise(count = n())
game$NUMFOULS <- temp[match(game$GROUP, temp$GROUP), "count"]
game$NUMFOULS[is.na(game$NUMFOULS)] <- 0
game$GROUP <- NULL

# Plot SCOREMARGIN vs TIME and NUMFOULS vs TIME
ggplot(game, aes(game$SECINTOGAME, game$SCOREMARGIN)) + geom_line(aes(colour = game$NUMFOULS)) + scale_x_continuous(breaks = c(0, 720, 1440, 2160, 2880), labels = c('P1', 'P2', 'P3', 'P4', 'Final')) + labs(title = 'Spurs vs Clippers', x = 'Period', y = 'Score Margin') + theme(legend.title=element_blank())

```