---
title: "Data_Processing"
author: "Team"
date: "April 18, 2016"
output: html_document
---
#EVENTMSGTYPE
1 - Make 2 - Miss 3 - Free Throw 4 - Rebound 5 - out of bounds / Turnover / Steal 6 - Personal Foul 7 - Violation 8 - Substitution 9 - Timeout 10 - Jumpball 12 - Start Q1? 13 - Start Q2?

#EVENTMSGACTIONTYPE
1 - Jumpshot 2 - Lost ball Turnover 3 - ? 4 - Traveling Turnover / Off Foul 5 - Layup 7 - Dunk 10 - Free throw 1-1 11 - Free throw 1-2 12 - Free throw 2-2 40 - out of bounds 41 - Block/Steal 42 - Driving Layup 50 - Running Dunk 52 - Alley Oop Dunk 55 - Hook Shot 57 - Driving Hook Shot 58 - Turnaround hook shot 66 - Jump Bank Shot 71 - Finger Roll Layup 72 - Putback Layup 108 - Cutting Dunk Shot



```{r setup, include=FALSE}
setwd("/Users/bobminnich/Documents/Columbia/Courses/Applied_Data_Science/FinalProject/finalproject-p5-team3")
library(dplyr)
library(ggplot2)
load("teamdata.Rdata")
process <- function(teamdata){
  teamdata$EVENTMSGTYPE = as.numeric(teamdata$EVENTMSGTYPE)
  teamdata$GAME_ID = as.factor(as.character(teamdata$GAME_ID))
  teamdata$EVENTMSGTYPE = as.numeric(teamdata$EVENTMSGTYPE)
  teamdata$HOMEDESCRIPTION = tolower(as.character(teamdata$HOMEDESCRIPTION))
  teamdata$VISITORDESCRIPTION = tolower(as.character(teamdata$VISITORDESCRIPTION))
  teamdata$PERIOD = as.numeric(teamdata$PERIOD)
  teamdata$EVENTNUM = (as.numeric(teamdata$EVENTNUM))
  teamdata$WCTIMESTRING = as.character(teamdata$WCTIMESTRING)
  teamdata$PCTIMESTRING = as.character(teamdata$PCTIMESTRING)
  teamdata$SCORE = as.character(teamdata$SCORE)
  
  teamdata$hscore = "NULL"
  teamdata$vscore = "NULL"
  
  teamdata$hscore <- as.numeric(unlist(lapply(strsplit(teamdata$SCORE, "\\-"), "[", 1)))
  teamdata$vscore <- as.numeric(unlist(lapply(strsplit(teamdata$SCORE, "-"), "[", 2)))

  teamdata = teamdata[!duplicated(teamdata),] 
  
  teamdata$SCOREMARGIN[1] = '0'
  for(i in 1:nrow(teamdata)) {
    if(teamdata$SCOREMARGIN[i] == 'TIE')
      teamdata$SCOREMARGIN[i] <- '0'
    else if(teamdata$SCOREMARGIN[i] == 'NULL')
      teamdata$SCOREMARGIN[i] <- teamdata$SCOREMARGIN[i - 1]
  }
  teamdata$SCOREMARGIN <- as.numeric(teamdata$SCOREMARGIN)
  
  # 
  # df_fouls = filter(teamdata, EVENTMSGTYPE=="6" ) %>% group_by(GAME_ID) %>% summarise(count = n())
  # #0021400020 had the most fouls Thunder vs Clippers
  # which.max(df_fouls$count)
  # ggplot(df_fouls, aes(x=count)) +
  #     geom_histogram(binwidth=1, alpha=.5, position="identity")
  # 
  
  
  
  teamdata$HFoulName = "NULL"
  teamdata$HMake = "NULL"
  teamdata$HFreeName = "NULL"
  teamdata$HTake = 0
  
  teamdata$VFoulName = "NULL"
  teamdata$VMake = "NULL"
  teamdata$VFreeName = "NULL"
  teamdata$VTake = 0
  
  teamdata$min <- lapply(strsplit(teamdata$PCTIMESTRING, ":"), "[", 1)
  teamdata$sec <- lapply(strsplit(teamdata$PCTIMESTRING, ":"), "[", 2)
  
  teamdata$min2 <- floor(as.numeric(teamdata$min)/2) 
  
  teamdata$Time <- (as.numeric(teamdata$PERIOD) - 1)*12*60 +abs((as.numeric(teamdata$min)-12)*60 - (0-as.numeric(teamdata$sec)))
  
  for(i in 1:nrow(teamdata)){
    if(grepl("foul", teamdata$HOMEDESCRIPTION[i])){
      if(grepl("take", teamdata$HOMEDESCRIPTION[i])){
        teamdata$HFoulName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[1]
        teamdata$HTake[i] =1
        teamdata$HMake[i] = "NULL"}
      else{
        teamdata$HFoulName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[1]
        teamdata$HMake[i] = "NULL"
        }
    }
      else if(grepl("miss", teamdata$HOMEDESCRIPTION[i]) & grepl("foul", teamdata$HOMEDESCRIPTION[i])){
      teamdata$HFreeName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[2]
      teamdata$HMake[i] = 0
  }
    else if(grepl("free", teamdata$HOMEDESCRIPTION[i])){
      teamdata$HFreeName[i] = unlist(strsplit(teamdata$HOMEDESCRIPTION[i],"\\s"))[1]
      teamdata$HMake[i] = 1
    }else{
      teamdata$HFreeName[i] = "NULL"
      teamdata$HFoulName[i] = "NULL"
      teamdata$HMake[i] = "NULL"
    }
    
    if(grepl("foul", teamdata$VISITORDESCRIPTION[i])){
      if(grepl("take", teamdata$VISITORDESCRIPTION[i])){
        teamdata$VFoulName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[1]
        teamdata$VTake[i] =1
        teamdata$VMake[i] = "NULL"}
      else{
        teamdata$VFoulName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[1]
        teamdata$VMake[i] = "NULL"
        }
    }
      else if(grepl("miss", teamdata$VISITORDESCRIPTION[i])){
      teamdata$VFreeName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[2]
      teamdata$VMake[i] = 0
  }
    else if(grepl("free", teamdata$VISITORDESCRIPTION[i])){
      teamdata$VFreeName[i] = unlist(strsplit(teamdata$VISITORDESCRIPTION[i],"\\s"))[1]
      teamdata$VMake[i] = 1
    }else{
      teamdata$VFreeName[i] = "NULL"
      teamdata$VFoulName[i] = "NULL"
      teamdata$VMake[i] = "NULL"
    }
  }
  
  teamdata$min4 <- floor(as.numeric(teamdata$min)/4)
  teamdata$min4r <-floor(((teamdata$Time)/60)/4) + 1
  teamdata$min4r[floor(teamdata$Time)==2880] = 12
  
  teamdata$SCORE = as.character(teamdata$SCORE)

  teamdata$hscore = "NULL"
  teamdata$vscore = "NULL"
  
  teamdata$hscore <- as.numeric(unlist(lapply(strsplit(teamdata$SCORE, "\\-"), "[", 1)))
  teamdata$vscore <- as.numeric(unlist(lapply(strsplit(teamdata$SCORE, "-"), "[", 2)))


return(teamdata)
  }
```

```{r}

hack_games = function(names,teamdata){
  df_game = filter(teamdata, EVENTMSGTYPE=="3",HFreeName%in%names | VFreeName%in%names )%>%group_by(GAME_ID)%>% summarise(count = n())
  df_game = df_game[complete.cases(df_game),]
  mean(df_game$count)
  median(df_game$count)
  range(df_game$count)
  quant = quantile(df_game$count)
  print(quant)
  a = ggplot(df_game, aes(x=count)) +
      geom_histogram(binwidth=.5, colour="red", fill="white") + 
    xlab("Number of Free Throws Per Game for Poor Shooting")+
    ylab("Number of Games")
  
  games = df_game[df_game$count>as.numeric(quantile(df_game$count)[4]),]$GAME_ID
  return (list(games,a))
}


```

```{r}
load("teamdata.Rdata")
teamdata = process(teamdata)
test1 = hack_games("jordan",teamdata)
plot_clippers = test1[2]
games_clippers = test1[1]
plot_clippers[1]



load("teamdata2.Rdata")
teamdata2 = process(teamdata2)
test2 = hack_games(c("howard","smith"),teamdata2)
plot_rockets = test2[2]
games_rockets = test2[1]
plot_rockets[1]

load("teamdata3.Rdata")
teamdata3 = process(teamdata3)
test3 = hack_games("drummond",teamdata3)
plot_detriot = test3[2]
games_detriot = test3[1]
plot_detriot[1]
```



#ML PORTION

```{r}
names = "drummond"

df_game = filter(teamdata3, EVENTMSGTYPE=="3",HFreeName%in%names | VFreeName%in%names )%>%group_by(GAME_ID)%>% summarise(count = n())
  df_game = df_game[complete.cases(df_game),]
  mean(df_game$count)
  median(df_game$count)
  range(df_game$count)
  quant = quantile(df_game$count)
  print(quant)
  a = ggplot(df_game, aes(x=count)) +
      geom_histogram(binwidth=.5, colour="red", fill="white") + 
    xlab("Number of Free Throws Per Game for Poor Shooting")+
    ylab("Number of Games")
  
  games = df_game[df_game$count>as.numeric(quantile(df_game$count)[4]),]$GAME_ID


```




```{r}
#By game and by Player Jordan
game_list = unique(teamdata$GAME_ID)
df = data.frame(matrix(0,(length(game_list)),12*2))


for(i in 1:length(game_list)){
  onegame = filter(teamdata, GAME_ID == game_list[i], HFreeName == "jordan" | VFreeName == "jordan")
  grouping = group_by(onegame, min4r)
  grouping2 = filter(grouping, EVENTMSGTYPE == "3")
  final = summarise(grouping2, count = n())
  df$game[i] = as.character(game_list[i])
  
  games
  if(df$game[i] %in% games){
    #df[i,1:12] = 0
    df$HACK[i] = 1
  }else{
    df$HACK[i] = 0
  }
  if(dim(final)[1]==0){
    #df[i,1:12] = 0
    print("EMPTY")
  }else{
    df[i,final$min4r] = final$count
  }
  onegame = filter(teamdata, GAME_ID == game_list[i], is.numeric(SCOREMARGIN), min4r<=12) %>%group_by(min4r)%>%summarise(avg = mean(SCOREMARGIN))
  
  if(dim(onegame)[1]==0){
    df[i,(onegame$min4r+12)] = 0
    print("EMPTY")
  }else{
    df[i,(onegame$min4r+12)] = onegame$avg
  }
  
}

```


```{r}
load("teamstatsData.Rdata")
load("playerstatsData.Rdata")
load("visitinghome.Rdata")

```

#Need to pull other teams and output games
```{r}

load("teamdata.Rdata")
teamdata = process(teamdata)
maximin = function(hscore){
  return (max(hscore) - min(hscore))
}

df_game = filter(teamdata, hscore!="NULL")%>%group_by(GAME_ID,min4r)%>% summarise(diff_h =maximin(hscore) )

a = filter(teamdata, hscore!="NULL")

```

#Need to combine other 
